---
title: "NSD & delta Elevation"
output: pdf_document
date: "`r Sys.Date()`"
params: 
  elk_data: "elk"
  dem_data: "elk_dem"
editor_options: 
  chunk_output_type: console
---

Note that net squared displacement is measured from the initial collaring location.

Delta elevation is calculated as an elk's current elevation minus the minimum elevation of that elk. Therefore, all elk are measured from a baseline elevation of 0 meters for the purposes of plotting. This is done to account for elk that live close to shore, and therefore have a lower baseline elevation, versus elk that live more in the interior, which has a higher baseline elevation, even when that elk is in a valley bottom.

NSD and delta elevation are both normalized to run from 0 to 1 for the plots where both NSD and elevation show up simultaneously, to ensure they both share the same y-axis. *This erases the magnitude of change.* Only use these plots to see if the pattern of movement (NSD vs delta elevation) lines up.

Finally, for the NSD vs Elevation plots, I don't know if these are useful at all, but I threw them in thinking points falling along a 1:1 line indicates that NSD and delta elevation are tightly correlated. 

```{r setup, include=FALSE}
library(ggplot2)
library(ggrastr) # to rasterize the point geoms, for faster pdf loading

knitr::opts_chunk$set(echo = FALSE)

elk <- sf::st_drop_geometry(params$elk_data)

elk_dem <- params$dem_data

elk <- merge(elk, elk_dem, by = "idposition")

elk <- elk |> 
  dplyr::group_by(animal_id, collar_id) |>
  dplyr::arrange(dttm) |>
  dplyr::mutate(min_elevation_m = min(elevation_m),
                d_elevation_m = elevation_m - min_elevation_m)

quad_plot <- function(elk, id) {
  # p1
  # dttm vs NSD
  p1 <- elk |>
    dplyr::filter(animal_id == id) |>
    dplyr::mutate(date = lubridate::date(dttm),
                  NSD = NSD / 1000 / 1000) |> # Convert NSD in meters to NSD in km
    ggplot(aes(x = date,
               y = NSD)) +
    geom_path(aes(group = as.factor(collar_id)),
              alpha = 0.3, 
              color = "black") +
    geom_point_rast(pch = ".",
                    color = "darkgreen") +
    ggplot2::scale_x_date(labels = scales::date_format("%b %y"),
                          date_breaks = "3 months") +
    labs(title = "NSD",
         y = "Net Squared Dislacement (km)") +
    theme_minimal() +
    ggplot2::theme(title = element_text(size = 8),
                   axis.title = element_text(size = 8),
                   axis.title.x = element_blank(),
                   axis.ticks.x = element_line(color = "black"),
                   axis.text.x = element_text(size = 7, angle = 90))

  # p2 
  # dttm vs delta elevation
  p2 <- elk |> 
    sf::st_drop_geometry() |>
    dplyr::group_by(animal_id) |>
    dplyr::arrange(dttm) |>
    dplyr::mutate(min_elevation_m = min(elevation_m),
                  d_elevation_m = elevation_m - min_elevation_m) |>
    dplyr::filter(animal_id == id) |>
    dplyr::mutate(date = lubridate::date(dttm)) |> 
    ggplot(aes(x = date,
               y = elevation_m)) +
    geom_path(aes(group = as.factor(collar_id)),
              alpha = 0.3, 
              color = "black") +
    geom_point_rast(pch = ".",
                    color = "orangered") +
    ggplot2::scale_x_date(labels = scales::date_format("%b %y"),
                          date_breaks = "3 months") +
    labs(title = "Elevation",
         y = "delta elevation (m)") +
    theme_minimal() +
    ggplot2::theme(title = element_text(size = 8),
                   axis.title = element_text(size = 8),
                   axis.title.x = element_blank(),
                   axis.ticks.x = element_line(color = "black"),
                   axis.text.x = element_text(size = 7, angle = 90))
  
  # p3
  # dttm vs both NSD & elevation
  p3 <- elk |> 
    sf::st_drop_geometry() |>
    dplyr::filter(animal_id == id) |> 
    dplyr::mutate(NSD = NSD / 1000 / 1000) |> # expess in km rather than m
    # Normalize both to run from 0 to 1
    dplyr::mutate(NSD = NSD / max(NSD),
                  d_elevation_m = d_elevation_m / max(d_elevation_m),
                  date = lubridate::date(dttm)) |>
    dplyr::select(animal_id, collar_id, date, NSD, d_elevation_m) |>
    tidyr::pivot_longer(cols = c(NSD, d_elevation_m),
                        values_to = "normalized_delta") |>
    ggplot(aes(x = date,
               y = normalized_delta)) +
    geom_path(aes(group = interaction(name, collar_id)),
              alpha = 0.3, 
              color = "black") +
    geom_point_rast(pch = ".",
                    aes(color = name)) +
    scale_color_manual(values = c("orangered", "darkgreen")) +
    ggplot2::scale_x_date(labels = scales::date_format("%b %y"),
                          date_breaks = "3 months") +
    labs(title = "Both NSD & Elevation",
         y = "Normalized Delta") +
    theme_minimal() +
    ggplot2::theme(title = element_text(size = 8),
                   axis.title = element_text(size = 8),
                   axis.title.x = element_blank(),
                   axis.ticks.x = element_line(color = "black"),
                   axis.text.x = element_text(size = 7, angle = 90),
                   legend.position = "none")
  
  p4 <- elk |> 
    sf::st_drop_geometry() |>
    dplyr::filter(animal_id == id) |> 
    dplyr::mutate(NSD = NSD / 1000 / 1000) |> # expess in km rather than m
    # Normalize both to run from 0 to 1
    dplyr::mutate(NSD = NSD / max(NSD),
                  d_elevation_m = d_elevation_m / max(d_elevation_m)) |>
    ggplot(aes(x = NSD, 
               y = d_elevation_m)) +
    geom_point_rast(pch = ".",
                    color = "slateblue") +
    geom_abline(slope = 1) +
    labs(title = "NSD vs Elevation",
         x = "Normalized NSD",
         y = "Normalized delta elevation (m)") +
    theme_minimal() +
    ggplot2::theme(title = element_text(size = 8),
                   axis.title = element_text(size = 8),
                   axis.ticks.x = element_line(color = "black"),
                   axis.text.x = element_text(size = 7, angle = 90))

  p <- ggpubr::ggarrange(p1, p2, p3, p4) |>
    ggpubr::annotate_figure(top = ggpubr::text_grob(id,
                                                    color = "black",
                                                    face = "bold",
                                                    size = 12))
  return(p)
}

ids <- unique(elk$animal_id)
```

```{r plots, fig.align='center', fig.height=4.2}
for (i in ids) {
  p <- suppressMessages(quad_plot(elk, i))
  print(p)
}
```